---
title: "Manipulating ABCD rsfmri output in R"
author: "Eric "Fez" Feczko"
date: "2024-11-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Raveio install

This will setup raveio, which is needed to load ABCD data into R easily. There are other tools that work as well, however, the organization of the outputs may be difficult to use. 

Raveio requires specific system dependencies:

libhdf5-dev (https://stackoverflow.com/questions/39236025/how-to-install-libhdf5-dev-without-yum-rpm-nor-apt-get)
fftw (`sudo apt-get install fftw3 fftw3-dev pkg-config`)

```{r raveio install}
install.packages("raveio", repos = c("https://rave-ieeg.r-universe.dev", "https://cloud.r-project.org"))
```

## opening .mat files using raveio and reformatting data into a datasheet

The raveio library contains input/output commands for reading matlab v 7.3 files. Below is an example of how to load the file:

```{r open mat file}
library(raveio)
path_to_mat_file <- "/home/feczk001/shared/projects/ABCD/abcd-sync/data/ABCD6p0_prerelease/rsfmri_091824/rsfmri/gpnet_corr.mat"
path_to_info_file <- "/home/feczk001/shared/projects/ABCD/abcd-sync/data/ABCD6p0_prerelease/rsfmri_091824/rsfmri/vol_info.mat"
gpnet_corr_matfile <- read_mat(path_to_mat_file)
vol_info_matfile <- read_mat(path_to_info_file)
```
## reorganizing data into a dataframe 

The matlab files may need reformatting before data analysis in R, the code below shows one how to create a data frame from the correlations and label it with the participant and session ids as well as the number of "good" quality frames (ntpts). 

```{r organize data into dataframe}
pid <-  unlist(vol_info_matfile$participant.id)
sid <- unlist(vol_info_matfile$session.id)

network_corr_label <- paste(gpnet_corr_matfile$net1vec,gpnet_corr_matfile$net2vec, sep = "_")
corrdata <- data.frame("participant" = pid,"session" = sid, "ntpts" = gpnet_corr_matfile$ntpointvec,gpnet_corr_matfile$corrmat)
column_labels <- c("participant","session","ntpts",network_corr_label)
colnames(corrdata) <- column_labels
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
